---
title: "STAT 345 Midterm Project"
font: 12pt
date: "Due April 4"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
```{r, echo = FALSE, warning=FALSE, message = FALSE}

library(devtools)
library(tidyverse)
library(nbastatR)
library(ggplot2)
library(dplyr)
library(purrr)
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2) #Fix to load in the NBA data without error


bucks_shots_24_unclean <- teams_shots(
  teams = "Milwaukee Bucks",
  team_ids = NULL,
  all_active_teams = F,
  season_types = "Regular Season",
  seasons = 2024,
  measures = "FGA",
  periods = 0,
  months = 0,
  date_from = NA,
  date_to = NA,
  nest_data = F,
  return_message = F
) #Loading in shooting data for the most recent full season

bucks_shots_24 <- bucks_shots_24_unclean %>% select("isShotMade", "locationX", "locationY", "yearSeason", "namePlayer") #Selecting only relevant columns: shot outcome, shot location, year of season, name of player 

bucks_shots_19_unclean <- teams_shots(
  teams = "Milwaukee Bucks",
  team_ids = NULL,
  all_active_teams = F,
  season_types = "Regular Season",
  seasons = 2019,
  measures = "FGA",
  periods = 0,
  months = 0,
  date_from = NA,
  date_to = NA,
  nest_data = F,
  return_message = F
) #Loading in data for 2019 season, five years from the most recent full season
bucks_shots_19 <- bucks_shots_19_unclean %>% select("isShotMade", "locationX", "locationY", "yearSeason", "namePlayer") #Selecting only relevant columns: shot outcome, shot location, year of season, name of player 

bucks_shots_14_unclean <- teams_shots(
  teams = "Milwaukee Bucks",
  team_ids = NULL,
  all_active_teams = F,
  season_types = "Regular Season",
  seasons = 2014,
  measures = "FGA",
  periods = 0,
  months = 0,
  date_from = NA,
  date_to = NA,
  nest_data = F,
  return_message = F
) #Loading in data for 2014 season, 10 years from the most recent full season
bucks_shots_14 <- bucks_shots_14_unclean %>% select("isShotMade", "locationX", "locationY", "yearSeason", "namePlayer") #Selecting only relevant columns: shot outcome, shot location, year of season, name of player 

bucks_shots_09_unclean <- teams_shots(
  teams = "Milwaukee Bucks",
  team_ids = NULL,
  all_active_teams = F,
  season_types = "Regular Season",
  seasons = 2009,
  measures = "FGA",
  periods = 0,
  months = 0,
  date_from = NA,
  date_to = NA,
  nest_data = F,
  return_message = F
) #Loading in data for 2009 season, 15 years from the most recent full season
bucks_shots_09 <- bucks_shots_09_unclean %>% select("isShotMade", "locationX", "locationY", "yearSeason", "namePlayer") #Selecting only relevant columns: shot outcome, shot location, year of season, name of player 


bucks_shots_list <- list(bucks_shots_24, bucks_shots_19, bucks_shots_14, bucks_shots_09)#combining all of the shot data from 2009, 2014, 2019, and 2024 into one list
#These years were used because I wanted data from every five years

bucks_shots_all <- reduce(bucks_shots_list, full_join, by = colnames(bucks_shots_24)) #reducing all of shot data that way it could be faceted by year as graphs
```
```{r, echo = FALSE, warning=FALSE, message = FALSE}
library(sportyR)
library(ggdark)

bball_court <- geom_basketball("nba", x_trans = 43, rotation = 90, display_range = "defense") #Transform the x axis by 43 to change the origin to be around the hoop 
#Rotate by 90 degrees and display range = defense to view the area of the court the data was plotted on

bball_court + 
geom_point(data = bucks_shots_all, size = .3, alpha = .2, #Size is smaller and opacity is used to make the points easier to read
           aes(locationX/10, locationY/10, color = isShotMade)) + #the X/Y coordinates are divided by 10 because the values represented a tenth of foot and the court is plotted based on feet
scale_colour_manual(values = c( "#8B0000", "#006400"), labels = c("Missed Shots", "Made Shots")) + #Colors based on whether the shot was made are not with green/red used because it is easier to understand at a glance

  
  labs(
    title = "Milwaukee Bucks Shoting Patterns Over Time",
    color = "Shot Outcome"
  ) + scale_x_reverse() + #The shot data is said to be inverse so this a fix so that it represents where the shots where actually at and not based on the view of the audience
facet_wrap(~ yearSeason) + #Facet by year that way the points in the graphs are less crowded and lends itself to better analysis and viewing
guides(color = guide_legend(override.aes = list(size = 4, alpha = 1))) + #Makes the points in the legend not opaque that way it's easier to read
 dark_theme_classic() + #Dark theme works well against the graphs, mainly for aesthetics
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank()) #Exclude any X/Y axis lines because it wouldn't be appropriate on top of the court image
```


Main Findings:

*Over time, shots have become more clustered around the three-point line and close to the basket.

*The use of the three-point shot has increased over the last 15 years.

*Lay-up shots are more common in 2014-2024 than in 2009.

*Accuracy in lay up shots has increased over time.

*Compared to 2009, the 2024 season saw less shots taken past the three-point line.

*Shots taken were more spread out in 2009 and 2014.

Futher Inferences:

Over time the Milwaukee Bucks shoting locations have become more strategic. Shots are commonly taken on the three-point line. This can be inferred due to the increase in points clustered around the three-point line in the most recent season compared to seasons past. In 2009 and 2014 for example the point density in the graphs around the three-point line is less than in 2019 and 2024. In more recent seasons less shots were taken outside of the three-point line. It seems that this could be a more strategic use of the three-point line.

In 2024, shots are commonly clustered around the basket, indicating a use of lay-up shots. 2009 saw less utilization of this type of shot. Overtime there seems to be an increase in the number of lay-ups taken due to the clustering of points. This seems to be a more strategic shot location due in part to the likelihood of success and the evidence of a clustering of made shots indicated by the green points.

Overall, the shots taken by the Milwaukee Bucks have become more efficient and strategic of high scoring and high likelihood of success. The increased utilization of the three-point line indicates a strategy to score higher points per shot. The increased use of the lay-up shot and those close to the basket could indicate a strategy in taking shots that have a high likelihood of success. The clustering of green points in the later seasons backs up this strategy showing that made shots commoningly happen close to the basket. This in all would indicate that the Bucks have improved as a team from 2009-2024.


\newpage

```{r, echo = FALSE, warning=FALSE, message = FALSE, fig.width = 10}
salaries <- suppressMessages(suppressWarnings(hoopshype_salaries(
  teams = NULL,
  all_teams = TRUE, 
  assume_player_opt_out = TRUE,
  assume_team_doesnt_exercise = TRUE,
  nest_data = FALSE,
  return_message = FALSE
))) %>% #Salaries for NBA players pulled
  filter(urlTeamSalaryHoopsHype == "https://hoopshype.com/salaries/milwaukee_bucks/") #including only the Milwaukee Bucks

salaries_clean <- salaries %>% filter(slugSeason == "2024-25") %>% #Including future contracted pay for the season directly after the shot data plotted
  select(namePlayer, amountContract) %>% #selecing only relevant columns
  mutate(
    salary_bracket = case_when(
      amountContract >= 20e6 ~ "High",    
      amountContract >= 5e6  ~ "Middle",  
      TRUE                  ~ "Low"       
    )
  ) %>% #Creating different levels of salaries 
  # 20 million and above indicates a High salary, Between 20 million and 5 million indicates a Middle salary, Below 5 million indicates a Low salary 
  #These numbers were used because they capture an accurate amount of players and shots
  
  select(-amountContract) #Getting rid of contract amount and only have the salary range values (high, middle, low)

bucks_salaries <- bucks_shots_24 %>%
  left_join(salaries_clean, by = "namePlayer") %>% filter(!is.na(salary_bracket)) %>% mutate(salary_bracket = factor(salary_bracket, levels = c("High", "Middle", "Low"))) #Joining by Name of player and fixing the order of salary levels 

bball_court + 
  geom_point(data = bucks_salaries, 
             aes(locationX/10, locationY/10, color = isShotMade), #Fixing the scale of X/Y coordinates
             size = 0.5, alpha = 0.5) + #Increasing size and decreasing opacity since there are less points per plot
  scale_colour_manual(values = c( "#8B0000", "#006400"), labels = c("Missed Shots", "Made Shots")) +
  labs(
    title = "Past Shoting Effectiveness (2023-2024) and Future Contracted Pay (2024-2025)",
    x = "X Location",
    y = "Y Location",
    color = "Shot Outcome"
  ) + 
  scale_x_reverse() + #inverting the x axis
  facet_wrap(~ yearSeason) + 
  guides(color = guide_legend(override.aes = list(size = 4, alpha = 1))) + 
  dark_theme_classic() +  
  theme(
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +  facet_wrap(~ salary_bracket) #facet by salary level for easy comparison

```

Future Contracted Pay and Shots Findings:

In the above graphic, shot effectiveness and location for the 2024 season is graphed. Whether or not the shot was made is indicated by color. On the three different versions the shots are separated based on future contracted pay for the 2025 season. High pay contains shots made by players with over 20 million contracted. Middle pay contains shots made by players with contracted pay between 20 million and 5 million dollars. Low pay contains shots made by players with less than 5 million dollars contracted for the 2025 season. The players that are getting paid more in the future took more shots in 2024. The number of shots taken decreases as pay decreases. This indicates that star players will get paid more than players who are not in the spotlight. There seems to be similar clusterings around the three point line in all three levels of pay indicating that all players may have a preference for the three point line as a shoting location. 


   
 

   - 

  
  
  


